BLUP anaysis

### 2020-04-30 R/rrBLUP_analysis_D2F.Rmd
Analysis for Var_P and heritability within sites and subpops for flowering

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rlang)
library(rrBLUP)
library(switchgrassGWAS)
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rlang)
library(rrBLUP)
library(switchgrassGWAS)
library(viridis)
library(cowplot)
```

@@

# Per site analysis

GR50, D2F, FL50, and CGDD_13C

```{r}
gdd_all <- tmp2 %>%
  group_by(SITE, PLOT_GL, PLANT_ID, FL_end, GR_start) %>%
  summarise(CGDD_0C = sum(GDD_0C_SM),
            CGDD_8C = sum(GDD_8C_SM),
            CGDD_13C = sum(GDD_13C_SM),
            CDYLN = sum(DYLN)
            ) %>%
  mutate(D2F = FL_end - GR_start) %>%
  left_join(metadata) %>%
  dplyr::select(-(LIBRARY:PLANTED_2018)) %>%
  left_join(phe_all)
```

```{r}
i=1; j = 7; k=3
(i-1)*length(fourway_site_list)*length(subpop_list) + (k-1)*length(subpop_list) + j
```
15 phe (i), 7 sites (k), 3 subpops (j)

Would like to compare subpops at each site and for each phe.
so same phe number is closest together

i*j*k + j*k + j

```{r}
PHE <- list(quote(CGDD_13C), quote(CGDD_8C), quote(CGDD_0C), quote(D2F), quote(FL50), quote(GR50), quote(CDYLN), quote(BIOMASS), quote(TC_EOS), 
            quote(HT_PAN_EOS), quote(FL1), quote(EM1), quote(EM50),
             quote(GR1))  # quote(TC_EOS_2018),
fourway_site_list <- list(quote(BRKG), quote(CLMB), quote(KBSM), quote(KING), quote(LINC), quote(PKLE), quote(STIL))
subpop_list <- list(Midwest_and_Gulf = c("Midwest", "Gulf"), Gulf = c("Gulf"),
                    Midwest = c("Midwest")) #, Full = c("Atlantic", "Gulf",
                                             #        "Midwest"))

rrkinblupout <- list()
kinseqblup <- list()
bluptable <- list()

for(i in seq_along(PHE)){
  for(k in seq_along(fourway_site_list)){
    for(j in seq_along(subpop_list)){

      list_counter <- (i-1)*length(fourway_site_list)*length(subpop_list) + (k-1)*length(subpop_list) + j
  testdf <- gdd_all %>%
    left_join(metadata, by = c("PLANT_ID", "SUBPOP")) %>%
    filter(!is.na(eval(PHE[[i]])) & 
             GENETIC_SUBPOP_KINSHIP %in% subpop_list[[j]] & 
             SITE %in% deparse(fourway_site_list[[k]])) %>%
    as.data.frame()

  bluprrk <- kin.blup(data = testdf, geno = "PLANT_ID", 
                      pheno = deparse(PHE[[i]]), K = V_g) #, fixed = c("SITE")

  rrkinblupout[[list_counter]] <- enframe(bluprrk$g, name = "PLANT_ID", 
                               value = deparse(PHE[[i]]))

    # Which Taxa were phenotyped for this phenotype? Just keep those for GWAS.
  taxa_phenotyped <- testdf %>%
    dplyr::select(PLANT_ID, !! PHE[[i]]) %>%
    as_tibble() %>%
    group_by(PLANT_ID) %>%
    summarise(count = n())
  kinseqblup[[list_counter]] <- taxa_phenotyped %>%
    left_join(rrkinblupout[[list_counter]], by = "PLANT_ID")

  bluptable[[list_counter]] <- tibble(Phenotype = deparse(PHE[[i]]), 
                                      V_a = bluprrk$Vg, V_e = bluprrk$Ve,
                                      SITE = deparse(fourway_site_list[[k]]),
                                      Subpop = names(subpop_list)[j]) %>%
    mutate(h2 = V_a / (V_a + V_e))

  #saveRDS(bluprrk, paste0("~/Github/pvdiv-phenology-gxe/data-raw/ignore/", 
  #                        "Kinship_BLUPs_", fourway_site_list[[k]], "_",
  #                        PHE[[i]], "_", nrow(taxa_phenotyped), "g_",
  #                        names(subpop_list)[j], ".rds"))
    }
  }
}

```{r}
BLUPs <- select(metadata, PLANT_ID, SUBPOP) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
  left_join(dplyr::select(kinseqblup[[1]], PLANT_ID, 3))

for(i in seq_along(bluptable)[-1]){
  if(!is.null(bluptable[[i]])){
  BLUPtable <- BLUPtable %>%
    full_join(bluptable[[i]])
  }
}

BLUPtable
gdd_all %>% left_join(metadata) %>%
  pivot_longer(values_to = "MEAS", cols = c(CGDD_0C:D2F, TC_EOS:GR50), names_to = "PHE") %>%
  dplyr::select(-LIBRARY, -(GENETIC_SUBPOP_KINSHIP:PLANTED_2018)) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
  group_by(PHE, SITE, SUBPOP) %>%
  summarise(X_bar = mean(MEAS, na.rm = TRUE),
            V_P = var(MEAS, na.rm = TRUE),
            I = V_P / X_bar^2, 
            CV_P = 100*sqrt(V_P) / X_bar) %>%
  #arrange(desc(I)) %>%
  full_join(BLUPtable, by = c("PHE" = "Phenotype", "SITE", "SUBPOP" = "Subpop")) %>%
  mutate(I_A = V_a / X_bar^2,
         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
         CV_A = 100 * sqrt(V_a) / X_bar,
         CV_E = 100 * sqrt(V_e) / X_bar) %>%
  filter(SITE %in% fourway_sites) %>%
  arrange(SUBPOP, PHE, SITE) %>%
  write_csv(., "~/Github/pvdiv-phenology-gxe/data-raw/ignore/Phenotypic_CVs_Individual_sites.csv")

gdd_all %>% left_join(metadata) %>%
  pivot_longer(values_to = "MEAS", cols = c(CGDD_0C:D2F, TC_EOS:GR50), names_to = "PHE") %>%
  dplyr::select(-LIBRARY, -(GENETIC_SUBPOP_KINSHIP:PLANTED_2018)) %>%
  filter(SUBPOP %in% c("Midwest", "Gulf")) %>%
  group_by(PHE, SITE) %>%
  summarise(X_bar = mean(MEAS, na.rm = TRUE),
            V_P = var(MEAS, na.rm = TRUE),
            I = V_P / X_bar^2, 
            CV_P = 100*sqrt(V_P) / X_bar) %>%
  #arrange(desc(I)) %>%
  full_join(BLUPtable, by = c("PHE" = "Phenotype", "SITE")) %>%
  mutate(I_A = V_a / X_bar^2,
         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
         CV_A = 100 * sqrt(V_a) / X_bar,
         CV_E = 100 * sqrt(V_e) / X_bar) %>%
  filter(SITE %in% fourway_sites) %>%
  arrange(Subpop, PHE, SITE) %>%
  write_csv(., "~/Github/pvdiv-phenology-gxe/data-raw/ignore/Phenotypic_CVs_Individual_sites_2.csv")
```

```{r}
h2table <- read_csv("~/Github/pvdiv-phenology-gxe/data-raw/ignore/Phenotypic_CVs_Individual_sites_Gulf_Midwest.csv")

h2table %>%
  mutate(I_A = V_a / X_bar^2,
         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
         CV_A = 100 * sqrt(V_a) / X_bar,
         CV_E = 100 * sqrt(V_e) / X_bar) %>%
  filter(PHE == "D2F") %>%
  ggplot(aes(x = SUBPOP, y = h2)) + 
  geom_bar(stat = "identity") + facet_wrap(~SITE) + ylim(c(0,1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "../analysis/Heritability_D2F_by_Site_and_Subpop.png", 
          plot = last_plot())

h2table %>%
  mutate(I_A = V_a / X_bar^2,
         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
         CV_A = 100 * sqrt(V_a) / X_bar,
         CV_E = 100 * sqrt(V_e) / X_bar) %>%
  filter(PHE == "FL50") %>%
  ggplot(aes(x = SUBPOP, y = h2)) + 
  geom_bar(stat = "identity") + facet_wrap(~SITE) + ylim(c(0,1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "../analysis/Heritability_FL50_by_Site_and_Subpop.png", 
          plot = last_plot())

h2table %>%
  mutate(I_A = V_a / X_bar^2,
         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
         CV_A = 100 * sqrt(V_a) / X_bar,
         CV_E = 100 * sqrt(V_e) / X_bar) %>%
  filter(PHE == "GR50") %>%
  ggplot(aes(x = SUBPOP, y = h2)) + 
  geom_bar(stat = "identity") + facet_wrap(~SITE) + ylim(c(0,1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
save_plot(filename = "../analysis/Heritability_GR50_by_Site_and_Subpop.png", 
          plot = last_plot())

h2table %>%
  mutate(I_A = V_a / X_bar^2,
         CV_R = 100 * sqrt(V_P - V_a) / X_bar, 
         CV_A = 100 * sqrt(V_a) / X_bar,
         CV_E = 100 * sqrt(V_e) / X_bar) %>%
  filter(PHE %in% c("CGDD_0C", "CGDD_8C", "CGDD_13C", "D2F", "FL50", "CDYLN", "FL1", "GR50")) %>%
  ggplot(aes(x = PHE, y = h2)) + 
  geom_point(aes(color = SUBPOP)) + facet_wrap(~SITE) + ylim(c(0,1)) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_color_viridis(discrete = TRUE, end = 0.8, direction = -1)
save_plot(filename = "../analysis/Heritability_Various_flowering_time_measures_by_Site_and_Subpop.png", 
          plot = last_plot())
```
